---
title: "Manuscript1"
author: "Hardy Griesbauer"
date: "20/02/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries
library(tidyverse)
library(dplR)
library(knitr)
library(treeclim)

# load data
load(here::here("data","fdi2020.RData"))

```

## TABLE: Chronology statistics
The chunk below is set not to run, as it will take a long time.  Instead, the table is saved into the 'data' folder and can be manipulated by the next chunk.

```{r eval=FALSE}
# Function to create statistics for chronologies
  # Set up parameters for detrending

spline.yrs=60

# Function to extract information for sites
  rwi.extract=function(site) {
    
    
  rwi.ext.subFunc<-function(crnType) {  
    
    # set spline window
    spline.yrs=60
    
    # Get crnType
    X<-fdi2020$ring
    
    # get site
    a=X[[crnType]][[site]] 
    
    # Subset for years less than 2010
    a=a[which(as.integer(rownames(a))<=2005),]
    
    # detrend and compute site stats
    a.stats=rwi.stats(detrend(a,method="Spline",nyrs=spline.yrs),prewhiten=TRUE)
    eps.stats=rwi.stats.running(detrend(a,method="Spline",nyrs=spline.yrs),prewhiten=TRUE,window.length=65)
    
    # have to transpose if only one row of data
    if(dim(eps.stats)[2]==1) eps.stats=as.data.frame(t(eps.stats))
    
    b.stats=rwl.stats(a)
    
    # return
    return(data.frame(site=site,
                      first=min(b.stats$first),
                      last=max(b.stats$last),
                      year=max(b.stats$year),
                      n.trees=a.stats$n.trees,
                      rbar=eps.stats$rbar.bt[nrow(eps.stats)],
                      eps=eps.stats$eps[nrow(eps.stats)]
           )
    )
    
  }
  
 # Run the subfunction
   ew<-rwi.ext.subFunc("ew") %>% 
    dplyr::select(Site=site,ew.rbar=rbar,ew.eps=eps)
  
  lw<-rwi.ext.subFunc("lw") %>% 
    dplyr::select(Site=site,lw.rbar=rbar,lw.eps=eps)
  
  # RETURN
   
    rwi.ext.subFunc("rw") %>% 
    dplyr::select(Site=site,First=first,
                  Last=last,Year=year,N.Trees=n.trees,rw.rbar=rbar,rw.eps=eps) %>% 
    inner_join(ew,by="Site") %>% 
    inner_join(lw,by="Site") %>% 
      return()
  }
    
  
  
# Create chronology statistics
  x<-
      lapply(names(fdi2020$ring$rw),rwi.extract) %>% 
      do.call(rbind.data.frame,.)  
  
  crnStats<-
    fdi2020 %>% 
     pluck("clim") %>% 
     pluck("normal") %>% 
    dplyr::select(Site,Latitude,Longitude,Elevation) %>% 
    left_join(x,by="Site") %>% 
    mutate(Site=stringr::str_to_title(Site))
 
  # # Save 
  save(crnStats,file=here::here("data","crnStatsTable.RData"))

```

The chunk below will run the code:

```{r}
# Load chronology statistics
load(here::here("data","crnStatsTable.RData"))

crnStats %>% 
  kable()

```

```{r}
# EPS filter
crnStats %>% 
  filter(ew.eps<0.85 | lw.eps<0.85 | rw.eps<0.85)


```

Note that Parsnip and Wap1 have latewood EPS slightly below the EPS threshold of 0.85.  They were kept in the dataset because their EPS values are close to 0.85 and the other chronologies (RW and EW) have EPS above 0.85.

## FIGURE: Site map
Create a map showing sites.
```{r}
library(sf)
library(ggplot2)
library(bcmaps)

# Create sites data
sites <-
  fdi2020 %>% 
  pluck("clim") %>% 
  pluck("normal") %>% 
  dplyr::select(Site,Longitude,Latitude) %>% 
  st_as_sf(coords=c("Longitude","Latitude"))

# Select cities to plot
cityBC<-
  bcmaps::bc_cities() %>% 
  filter(NAME%in%c("Prince George", "Vancouver","Williams Lake","Kelowna"))


# Set CRS
st_crs(sites) = 4326

# Create site map
siteMap<-
  ggplot()+
  geom_sf(data=bc_neighbours(),mapping=aes(fill=name))+
  geom_sf(data=sites,mapping = aes(fill = Site)) + 
  geom_sf(data=cityBC,color="red",shape=24,fill="red",size=1.5)+
    theme(legend.position = "none")+
  geom_sf_label(data=cityBC,aes(label = NAME),
                fill=NA,
                nudge_x=-25000,
                nudge_y=-4000,
                label.size=0)+
  coord_sf(xlim = c(1020000,1280000), ylim = c(400000,1130000), expand = FALSE,crs=3005)+
    theme(legend.position = "none")

  siteMap

```

Can edit the map later.

## Figure: boxplot of correlations between different chronology types

*Note that I set this chunk not to run, because we only need to do this once*
```{r eval=FALSE}

# function to put sites together and compute correlation coefficients
crnExt<-function(site) {
  
  
  ew<-
    fdi2020$crn$ew %>% 
    rownames_to_column("year") %>% 
    .[,c("year",site)] %>% 
    setNames(c("year",paste(site,"ew",sep=".")))
  
  lw<-
    fdi2020$crn$lw %>% 
    rownames_to_column("year") %>% 
    .[,c("year",site)] %>% 
    setNames(c("year",paste(site,"lw",sep=".")))
  
 
    fdi2020$crn$rw %>% 
    rownames_to_column("year") %>% 
    .[,c("year",site)] %>% 
    setNames(c("year",paste(site,"rw",sep="."))) %>% 
    inner_join(ew,by="year") %>% 
    inner_join(lw,by="year") %>% 
    filter(year>=1940) %>%  # filter for 1940 time period
    dplyr::select(-year) %>% 
    cor() %>% # correlation matrix
    .[lower.tri(.)] %>% 
    setNames(c("rw.ew","rw.lw","ew.lw")) %>% 
          c(site=site,.) %>% 
      return()
    
}

# Now run for all sites
# and save to .RData file

crnCorr<-
  lapply(sites$Site,crnExt) %>% 
  do.call(rbind.data.frame,.) %>% 
  mutate_at(vars(2:4),function(x) as.numeric(as.character(x))) %>% # reformat to numeric
  setNames(c("Site","rw.ew","rw.lw","ew.lw"))


  save(crnCorr,file=here::here("data","crnCorrTable.RData"))
  

```

The following chunk will load the correlation matrix between sites (1940-2005) and create a table.

```{r}
# Load chronology statistics
load(here::here("data","crnCorrTable.RData"))

crnCorr %>% 
  kable(digits=2)

```

### Figure: Boxplot of correlations 
```{r}
crnCorr %>% 
  pivot_longer(-Site,names_to="crn",values_to="corr") %>% 
  ggplot()+
  aes(y=corr,fill=crn)+
  geom_boxplot()

```

We can see that correlations between earlywood and latewood are weakest.

### Figure: Ew-Lw correlations by site climate

```{r}
crnCorr %>% 
  left_join(fdi2020$clim$normal,by="Site") %>% 
  ggplot()+
  aes(y=ew.lw,x=AHM)+
  geom_point()

```

Note very strong relationship between Ew-LW correlations and AHM, MAP, SHM, etc...

There is one chronology that is an outlier, let's identify it.

```{r}
crnCorr %>% 
  left_join(fdi2020$clim$normal,by="Site") %>% 
  filter(AHM>25 & AHM<29) 


```

## Correlation and response coefficients
Calculate response and correlation coefficients, as per the following algorithm:

1. Use seasCorr() to identify important climatic seasons
2. Use correlation and response function analysis

#### seasCorr function
We run the seasCorr function on each site to identify seasonal windows with maximal correlations with growth.  

*Important: the climPrep() function is used to prepare climate data for seascorr.  This function will add very small random numbers to climate variables with SD==0, just so that seascorr will function.  This is ok for variable screening, as per seascorr, but we will use dcc() function with original climate data to derive correlation and response coefficients!*

*Also note that the below chunk is not set to run (takes a very long time)*

```{r eval=FALSE}

x<-lapply(names(fdi2020$crn$rw),seasCorrSite)

seasCorrCoef<-
  do.call(rbind.data.frame,x) %>% 
  mutate(site=str_sub(crn,1,-4)) %>% 
  mutate(crn=str_sub(crn,-2,-1)) %>% 
  dplyr::select(site,crn,climVar,window,end.month,coef,sig)

save(seasCorrCoef,file=here::here("data","seasCorrCoef.RData"))

```

We can load the correlation coefficients from seascorr here: 
```{r}

load(here::here("data","seasCorrCoef.RData"))

```





